---
title: "ASAP_data_prep"
author: "C. Savonen, CCDL for ALSF"
date: "12/6/2018"
output: html_document
---

## Purpose: Prep data for use in [ASAP online](https://asap.epfl.ch/)
```{r Import data}
# Import gene expression matrix data
salmon.data <- readRDS(file.path("data", "salmon.data.RDS"))
```

### Import metadata
```{r Set up metadata}
if (file.exists(file.path("data", "meta_data.csv"))) {
# Read in csv if it's been made, if not read in from GEOquery
  geo.meta <- read.csv(file.path("data", "meta_data.csv"), stringsAsFactors = FALSE)
} else {
  geo.meta <- GEOquery::getGEO("GSE84465", destdir = "data")
  geo.meta <- data.frame(geo.meta[[1]]@phenoData@data)
}
```

### Make batch file for ASAP to correct batches by.
In this case we will use the plate ids. 
ASAP wants the first column to be the sample IDs that match your sample data file
and the second column to be the corresponding batch the sample belongs to.
```{r Make batch info}
batch.info <- cbind(geo.meta$geo_accession, geo.meta$plate.id.ch1)
head(batch.info)
```

### Make sure the samples and the metadata are in the same order.
```{r Order samples}
# Put samples in same order 
geo.meta <- geo.meta[match(colnames(salmon.data)[-c(1:2)], geo.meta$geo_accession), ]
```

### ASAP doesn't want extra gene columns and only likes round numbers for the counts
Here we have salmon.data have it's genes as it's rownames, so even though we are 
deleting these columns, the information we need is still there.
```{r Round numbers}
# Format the data so ASAP will like it. 
salmon.data <- apply(salmon.data[,-c(1:2)], 2, round)
head(salmon.data[,1:10])
```

### Write data file in the format ASAP will take.
ASAP wants a typical gene by column gene expression matrix with samples with 
the column names and the first column being the gene names.
You can use other forms of separation other than "/t" but will need to specify
what you have used in a drop down menu on ASAP. 
```{r Write data file}
# Write these data to tab delimited text files for use in ASAP
write.table(salmon.data, file.path("data", "salmon.data.asap.txt"), sep = "\t",
            row.names = FALSE, quote = FALSE)
```

### Write batch file in same type of format
```{r Write batch file}
# Write batch info file for ASAP use
write.table(batch.info, file.path("data", "batch.info.txt"), row.names = FALSE, 
            col.names = FALSE, sep = "\t", quote = FALSE)
```

```{r}
sessionInfo()
```
