---
title: "Darmanis Normalization Evaluation"
author: "C. Savonen CCDL for ALSF"
date: "12/17/2018"
output: html_document
---
### Purpose: compare normalization methods for Darmanis et al data
```{r Get packages }
# Import functions for this analysis
source(file.path("scripts", "util", "clustering_statistics_functions.R"))

# ggplot2 library
library(ggplot2)

# These are the necessary packages
packages <- c("Rtsne", "NMI", "caret", "umap")

# Check if these packages are installed and install them if they aren't
lapply(packages, function(package) {
  if (!(package %in% installed.packages())) {
    install.packages(package)
  }
})
```

Import data that was previously obtained from submitting counts that were formatted
by `darmanis_asap_data_prep.Rmd` to [ASAP](https://asap.epfl.ch/)
```{r Import data}
# Get the file names of all the normalized files
normalized.files <- dir(file.path("darmanis_data", "normalized_darmanis"))

# Read in each of the normalization files
normalized.data <- lapply(normalized.files, function(file) {
                          readr::read_tsv(file.path("darmanis_data",
                                                    "normalized_darmanis",
                                                    file))
                          })

# Keep all the gene lists
genes <- lapply(normalized.data, function(x) x[,1])

# Keep the names with it.
names(normalized.data) <- gsub("\\.tab|darmanis_", "", normalized.files)
```

```{r Metadata Set Up}
# Read in the data
meta <- readr::read_tsv(file.path("darmanis_data", "GSE84465_meta.tsv"))
```

Retrieve cell type and plate batch for these samples:
```{r Retrieve cell type and plate batch for these samples}
# Get sample names from columns
samples <- colnames(normalized.data[[1]])[-1]

# Keep metadata only for the samples we have
meta <- meta[match(samples, meta$geo_accession), ]

# Cell types info as it's own vector
cell.types <- as.factor(meta$`cell type:ch1`)

# Plate batch info as a factor. Remove "plate id: " part.
plate.batch <- as.factor(gsub("plate id: ", "", meta$`characteristics_ch1.1`))
```

Run tsne clustering for each dataset:
```{r Make clusters with tsne}
# Run tsne on each dataset and extract
tsne <- lapply(normalized.data, function(dataset) {
               # Run tsne
               tsne.res <- Rtsne::Rtsne(t(dataset[,-1]), check_duplicates = FALSE)
               
               # Only keep the dimension coordinates
               tsne.res <- tsne.res$Y
               
               # Keep the sample names
               names(tsne.res) <- samples
                        
               return(tsne.res)
               })
```

Plot all the tsne's for all the datasets with labeling for cell type. 
```{r Plot the cell type data}
# Plot with cell type labels
cell.type.plots <- lapply(tsne, function(dataset) {
                          # Get data normalizaion name
                          set.name <- names(tsne)[parent.frame()$i[]]
                          
                          # Make plots for cell type and plate batch 
                          DimPlot(dataset, cell.types, xlabel = "tsne dim 1",
                                  ylabel = "tsne dim 2", name = set.name)
                    })

# Extract the legend
legend <- cowplot::get_legend(cell.type.plots[[1]])

# Surpressing the legend in the files
cell.type.plots <- lapply(cell.type.plots, function(a.plot) {
                          a.plot + theme(legend.position = 'none')
                          })

# Plot all plots and legend together
main.plot <- cowplot::plot_grid(
             cowplot::plot_grid(plotlist = cell.type.plots, ncol = 2, align='hv'),
             cowplot::plot_grid(NULL, legend, ncol = 2),
                                rel_widths=c(1, 0.2))

# Save to png
ggplot2::ggsave(main.plot, filename = "darmanis_cell_type_tsne.png")

# Print out plot here
main.plot
```

Plot the tsne data with batch info.
```{r Plot the batch data}
# Plot with plate batch labels
batch.plots <- lapply(tsne, function(dataset) {
                      # Get data normalizaion name
                      set.name <- names(tsne)[parent.frame()$i[]]
                          
                      # Make plots for cell type and plate batch 
                      DimPlot(dataset, plate.batch, xlabel = "tsne dim 1",
                              ylabel = "tsne dim 2", name = set.name)
                    })

# Plot in grid
main.plot <- cowplot::plot_grid(plotlist = batch.plots, align = "hv", ncol = 2)

# Save to png
ggplot2::ggsave(main.plot, filename = "darmanis_batch_tsne.png")

# Print out plot here
main.plot
```

Run cell type analyses on all the tsne datasets
```{r Run cell type analyses}
# Get knn and kmeans results for all tsne's of all datasets
cell.type.results <- lapply(tsne, function(dataset) {
                            # Get KNN kappa results
                            knn.results <- KnnEval(dataset, metadata = cell.types)
                            
                            # Get kmeans nmi and ari results
                            kmeans.results <- KmeansEval(dataset,
                                                         metadata = cell.types)
                            
                            # Put together into one dataframe
                            data.frame(knn.results, kmeans.results)
                            })
```

Run plate batch analyses on all the tsne datasets
Currently, I can't get KNN to work for the batch info on this dataset. 
```{r Run batch analyses}
# Get knn and kmeans results for all tsne's of all datasets
batch.results <- lapply(tsne, function(dataset) {
                        # Kmeans evaluation 
                        KmeansEval(dataset, metadata = plate.batch)
                        })
```

Plot all the cell type statistics on a boxplot
```{r Print the cell type plots}
# Melt the  list into one data.frame
cell.type.df <- reshape::melt(cell.type.results)

# Make the plot
celltype.plot <- ggplot(data = cell.type.df, aes(x = L1, y = value, fill = variable)) +
    geom_boxplot(position = position_dodge()) + 
    xlab("Normalization method") +
    facet_wrap(~variable)

# Save plots to png
ggplot2::ggsave(plot = celltype.plot, "darmanis_cell_type_stats_plots.png", width = 10)

# Print plot
celltype.plot
```

Plot all the batch statistics on a boxplot
```{r Print the batch plots}
# Melt the  list into one data.frame
batch.df <- reshape::melt(batch.results)

# Make the plot
batch.plot <- ggplot(data = batch.df, aes(x = L1, y = value, fill = variable)) +
    geom_boxplot(position = position_dodge()) + 
    xlab("Normalization method") +
    facet_wrap(~variable)

# Save plots to png
ggplot2::ggsave(plot = batch.plot, "darmanis_batch_stats_plots.png", width = 10)

# Print plot
batch.plot
```

Session info: 
```{r}
sessionInfo()
```


