---
title: "ASAP_data_prep"
author: "C. Savonen, CCDL for ALSF"
date: "12/6/2018"
output: html_document
---

## Purpose: Prep data for use in [ASAP online](https://asap.epfl.ch/)
```{r Import functions}
source(file.path("scripts", "asap_prep_functions.R"))
```

```{r Import data}
# Magrittr pipe
`%>%` <- dplyr::`%>%`

# Import gene expression matrix data
data <- readr::read_tsv(file.path("darmanis_data", "salmon_quants",
                                  "darmanis_data_counts.tsv"))
# Store gene names separately
gene <- data$gene
```

### Import metadata
```{r Set up metadata}
if (file.exists(file.path("darmanis_data", "meta_data.csv"))) {
# Read in csv if it's been made, if not read in from GEOquery
  geo.meta <- read.csv(file.path("darmanis_data", "meta_data.csv"), 
                       stringsAsFactors = FALSE)
} else {
  geo.meta <- GEOquery::getGEO("GSE84465", destdir = "darmanis_data")
  geo.meta <- data.frame(geo.meta[[1]]@phenoData@data)
}
```

### Make sure the samples and the metadata are in the same order.
```{r Order samples}
# Put samples in same order 
id.key <- readRDS(file.path("darmanis_data", "sample_id_key.RDS"))

# Make a sample conversion key
convert.key <- as.list(as.character(id.key$gsm.ids))
names(convert.key) <- as.character(id.key$run)

# Obtain GSM ids using conversion key and make these the column names 
colnames(data) <- dplyr::recode(colnames(data), !!!convert.key)

# Only keep metadata for samples in the set
geo.meta <- geo.meta[match(colnames(data)[-1], geo.meta$geo_accession), ]
```

### Make batch file for ASAP to correct batches by.
In this case we will use the plate ids. 
ASAP wants the first column to be the sample IDs that match your sample data file
and the second column to be the corresponding batch the sample belongs to.
```{r Make batch info}
# Make batch info matrix
batch.info <- cbind(geo.meta$geo_accession, geo.meta$plate.id.ch1)

# Let's exclude batches of "1"
batch.sum <- summary(as.factor(batch.info[,2]))
batch.one <- names(batch.sum)[which(batch.sum <= 2)]

# Make a regex statement to exclude these batches of one
regex <- paste0(paste0("^", batch.one, "$"), collapse = "|")
batch.info <- batch.info[ -grep(regex, batch.info[, 2]), ]

# Exclude these samples from the data
data <- data %>% dplyr::select(batch.info[,1])

# Look at a preview of what this looks like
head(batch.info)
```

### ASAP doesn't want extra gene columns and only likes round numbers for the counts
Here we have data have it's genes as it's rownames, so even though we are 
deleting these columns, the information we need is still there.
Filter out genes that don't have at least one count in 1% of cells and filter out 
samples that don't express at least 100 genes.
```{r Filter out genes}
data <- AsapFilter(data, genes, min_counts = 1, perc_genes = 0.01, num_genes = 100)
```

### Write data file in the format ASAP will take.
ASAP wants a typical gene by column gene expression matrix with samples with 
the column names and the first column being the gene names.
You can use other forms of separation other than "/t" but will need to specify
what you have used in a drop down menu on ASAP. 
```{r Write data file}
# Write these data to tab delimited text files for use in ASAP
readr::write_delim(data, file.path("darmanis_data", "asap_format.txt"))

# Write a copy of the counts in the normalized data file
readr::write_tsv(data, file.path("darmanis_data", "normalized_darmanis",
                                 "darmanis_counts.tab"))

```

### Write batch file in same type of format
```{r Write batch file}
# Write batch info file for ASAP use
readr::write_delim(data.frame(batch.info), 
                   file.path("darmanis_data", "batch.info.txt"))
```

Session info:
```{r}
sessionInfo()
```