---
title: "Filter Cutoff Investigation"
author: "C. Savonen, CCDL for ALSF"
date: "3/5/2019"
output: html_document
---

#### Purpose: Set up the tabula muris 10X dataset we are working with for normalization and subsequent steps. 

Subgoal: The cutoffs used for the smart-seq2 single cell data do not work well 
on the 10X single cell data. This is not surprising, but let's investigate the 
10X tab mur data so we can come up with some better cutoffs. 

Other note: For the purposes of saving room and making a more manageable dataset 
to work with, I am only using P4, and have excluded the Liver and spleen samples. 

```{r}
# Load ggplot2 library
library(ggplot2)

# magrittr pipe
`%>%` <- dplyr::`%>%`
```

Set up the data
```{r}
# Read in data
all.data <- readRDS(file.path("tab_mur_data", "counts_tab_mur.RDS"))

# Keep the gene names separate
gene.names <- rownames(all.data)
```
 
Make cell info dataframe for labeling purposes
```{r}
# Clean up sample names
colnames(all.data) <- gsub("_possorted_genome.bam", "", colnames(all.data))

# Make a dataframe of this cell info
cell.info <- data.frame(strsplit(colnames(all.data), "\\.")[-1]) %>% 
  t() 

# Change column names
colnames(cell.info) <- c("cell_barcode", "sample.id")

# Keep the full id names for each as well for matching purposes
cell.info <- data.frame(full.id = colnames(all.data), cell.info)
```

Read in the metadata file obtained from figshare, clean it up so we can merge it
with the sample names obtained in the cell.info dataframe. 

```{r}
# Import metadata file and make it with only the variables and samples we want
metadata <- readr::read_csv(file.path("tab_mur_data", "metadata_droplet.csv")) %>% 
  dplyr::select("channel", "mouse.id", "tissue") %>% 
  dplyr::filter(grepl("10X_P4", channel), tissue != "Liver", tissue != "Spleen") %>% 

# Make column name matching so we can use inner_join to merge the info
colnames(metadata)[1] <- "sample.id"
```

The liver samples had 95% of their data dropped in Alevin due to noisy cell 
barcodes. 
Spleen is excluded because then it became the only tissue that wasn't represented
in both mice. 

```{r}
# Merge these two frames together
metadata <- dplyr::inner_join(cell.info, metadata, by = "sample.id") %>%
  readr::write_tsv("metadata.tsv")
```
Make a data matrix for ease of collecting stats

```{r}
data.mat <- all.data %>% as.matrix()
```

For least stringent filter, let's determine what the smallest expression value is

```{r}
# Determine a detection cutoff
min.cutoff <- min(data.mat[which(data.mat!= 0)])
```

Let's make a detection matrix to just hold each data point as expressed or not

```{r}
# Make a matrix of detected or not detected logical values
boolean <- data.mat >= min.cutoff
```

Determine cutoff for how many samples should express a gene 

```{r}  
# Make a vector that has the number of samples that express a gene
gene.sum <- apply(boolean, 1, sum)

# Plot this information
plot(density(gene.sum), xlim = c(0, 300));
abline(v = 3)
```

Determine cutoff for how many genes should be expressed in a sample 

```{r}
# Make vector that has the number of genes expressed by each cell
sample.sum <- apply(boolean, 2, sum) 

# Put this with the cell.info so we can look at the distributions by samples
sample.sum.df <- data.frame(sample.sum, cell.info) 

# Plot with each sample as its own line
ggplot(sample.sum.df, aes(x = sample.sum, color = sample.id)) +
  geom_density() +
  geom_vline(aes(xintercept = 100))
```

Use cutoffs determined by looking at the plots above and filter the data matrix,
tack on gene names, write to tsv file.

```{r}
all.data %>% dplyr::select(gene.sum > 3) %>% 
  dplyr::filter(sample.sum > 50) %>% 
  dplyr::mutate("genes" = gene.names) %>% 
  readr::write_tsv("filtered_count_matrix_tab_mur.tsv")
```

```{r}
sessionInfo()
```
